Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-02-26 paper/main.tex
progress in section 3

Now we describe the construction of QBF formulas in detail. The
construction is parametrized by two integers $C,K\ge 0$, where $K$ is
the desired number of acceptance marks and $C$ is the number of
clauses of $\psi$. For simplicity, we assume that the reduced
automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$.
%Further, we consider only subsets $T\subseteq\bigcup_{\textrm{SCC}~S}\delta_S$
First we describe Boolean variables used in the constructed QBF
formulas.
\begin{itemize}
\item For each transition $t\in\delta$, variable $e_t$ says whether
  $t$ is in the current set $T$ or not.
  \[e_t=\left\{
      \begin{array}{ll}
        1 & \textrm{if }t\in T\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]
%  In fact, we consider variables 
\item For each transition $t\in\delta$ and acceptance mark $k\in N_K$,
  variable $n_{t,k}$ says whether $k$ is on the transition
  $\rem(t)$ or not.
  \[n_{t,k}=\left\{
      \begin{array}{ll}
        1 & \textrm{if }k\in\mks(\rem(t))\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]
\item For each $c\in\{1,2,\ldots,C\}$ and acceptance mark $k\in N_K$,
  variables $i_{c,k}$ and $f_{c,k}$ say whether the $c^\textrm{th}$
  clause of $\psi$ contains terms $\Inf k$ or $\Fin k$, respectively.
  \[i_{c,k}=\left\{
      \begin{array}{ll}
        1 & \textrm{if $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]
  \[f_{c,k}=\left\{
      \begin{array}{ll}
        1 & \textrm{if $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]

@@

It remains to define the subformula $\mcycle(\vec{e})$ with the intended meaning
\[
  \vec{e}\textrm{ represents a cycle} \iff \mcycle(\vec{e}).
\]
In fact, our reduction methods remains correct whenever ``implies'' holds.
If $\mcycle(\vec{e})$

  $\vec{e}$ represents a cycle. In fact, $\mcycle(\vec{e})$ needs to be
satisfied by all cycles, but our reduction technique remains correct
even if it is satisfied also by some sets $T\subseteq\delta$ that are
not cycles. The only consequence is that

Note that our reduction method remains
correct even if $\mcycle(\vec{e})$ isAs we will see, However, our reduction method remains
correct.

The intended meaning
of this formula is We offer three definitions:

\todo{Jan is here!}

\subsection{Lightweight version of $\mcycle(\vec{e})$ (level 1)}

\subsection{Balanced version of $\mcycle(\vec{e})$ (level 2)}

\subsection{Strict version of  $\mcycle(\vec{e})$ (level 3)}

>>>>

Now we describe the construction of QBF formulas in detail. The
construction is parametrized by two integers $C,K\ge 0$, where $K$ is
the desired number of acceptance marks and $C$ is the number of
clauses of $\psi$. For simplicity, we assume that the reduced
automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$.
%Further, we consider only subsets $T\subseteq\bigcup_{\textrm{SCC}~S}\delta_S$
First we describe Boolean variables used in the constructed QBF
formulas.
\begin{itemize}
\item For each transition $t\in\delta$, variable $e_t$ says whether
  $t$ is in the current set $T$ or not.
  \[e_t=\left\{
      \begin{array}{ll}
        1 & \textrm{if }t\in T\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]
%  In fact, we consider variables 
\item For each transition $t\in\delta$ and acceptance mark $k\in N_K$,
  variable $n_{t,k}$ says whether $k$ is on the transition
  $\rem(t)$ or not.
  \[n_{t,k}=\left\{
      \begin{array}{ll}
        1 & \textrm{if }k\in\mks(\rem(t))\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]
\item For each $c\in\{1,2,\ldots,C\}$ and acceptance mark $k\in N_K$,
  variables $i_{c,k}$ and $f_{c,k}$ say whether the $c^\textrm{th}$
  clause of $\psi$ contains terms $\Inf k$ or $\Fin k$, respectively.
  \[i_{c,k}=\left\{
      \begin{array}{ll}
        1 & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]
  \[f_{c,k}=\left\{
      \begin{array}{ll}
        1 & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\
        0 & \textrm{otherwise}    
      \end{array}\right.
  \]

@@

It remains to define the subformula $\mcycle(\vec{e})$ with the intended meaning
\[
  \vec{e}\textrm{ represents a cycle} \iff \mcycle(\vec{e}).
\]
In fact, only the direction ``$\Longrightarrow$'' is required for
correctness of our reduction method. If $\mcycle(\vec{e})$ holds for
some valuation of $\vec{e}$ that does not represent a cycle, then the
required equivalence
$\msat_\varphi(\vec{e}) \iff
\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ on this valuation of
$\vec{e}$ adds a superfluous constraint on $\psi$ represented by
$\vec{i},\vec{f}$ and $\rem$ represented by $\vec{n}$. These
superfluous constraints can lead to loss of reduction opportunities,
but not to incorrectness. This observation allows us to trade the
precision of $\mcycle(\vec{e})$ for simplicity of the formula.

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
  only minimal requirements on $\vec{e}$. It does not use the
  information about automaton structure, but it comes with an
  interesting simplification of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is a balanced version. It uses the
  information about the automaton structure to identify many
  valuations of $\vec{e}$ that are not cycles. It is relatively
  simple, but it is still satisfied by some valuations of $\vec{e}$
  not corresponding to cycles.
\item $\mcycle_3(\vec{e})$ is a strict version satisfied only by
  the valuations of $\vec{e}$ representing cycles. Unfortunately,
  it uses additional quantified variables. 
\end{itemize}

\subsection{Lightweight version $\mcycle_1(\vec{e})$}
The lightweight version is based on the fact that every cycle is nonempty.
Indeed, the only restriction imposed by
\[
  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
\]
is that $\vec{e}$ represents a nonempty set.

The fact that this version do

\subsection{Balanced version $\mcycle_2(\vec{e})$}

\subsection{Strict version $\mcycle_3(\vec{e})$}
